---
title: "R Notebook Conexión a base de datos"
output: html_notebook
---


#Conexión a bases de datos con información electoral de Chile

## Herramientas

```{r}
install.packages("ggplot2")
library(ggplot2)
```


## Probando paquete Car con datos Chile

https://www.rdocumentation.org/packages/car/versions/2.1-6/topics/Chile
"Voting Intentions in the 1988 Chilean Plebiscite

The Chile data frame has 2700 rows and 8 columns. The data are from a national survey conducted in April and May of 1988 by FLACSO/Chile. There are some missing data.

###Format

This data frame contains the following columns:

region
    A factor with levels: C, Central; M, Metropolitan Santiago area; N, North; S, South; SA, city of Santiago.
population
    Population size of respondent's community.
sex
    A factor with levels: F, female; M, male.
age
    in years.
education
    A factor with levels (note: out of order): P, Primary; PS, Post-secondary; S, Secondary.
income
    Monthly income, in Pesos.
statusquo
    Scale of support for the status-quo.
vote
    a factor with levels: A, will abstain; N, will vote no (against Pinochet); U, undecided; Y, will vote yes (for Pinochet).

###References

Fox, J. (2008) Applied Regression Analysis and Generalized Linear Models, Second Edition. Sage.
Fox, J. and Weisberg, S. (2011) An R Companion to Applied Regression, Second Edit

```{r}
install.packages("car")
library(car)
```





```{r}
ggplot(Chile, aes(statusquo, income, colour = vote)) + 
  geom_point()

ggplot(Chile, aes(statusquo, age , colour = region)) + 
  geom_point()


```

## Probando datos DataChile 

El sitio datachile.io tiene información pública del SERVEL
2009-2013-2016-2017	Electoral service (SERVEL)

lea: DataChile/readme.md


```{r }
#getwd()
#SelWorkingDir <- choose.dir()
#SelWorkingDir
file1<-"DataChile\\s2009-procesado.csv"
s2009_procesado<- read.csv(file=file1, header=TRUE, sep=",")
head(s2009_procesado)
```




## Probando paquete OECD 

The `OECD` package allows the user to download data from the OECD's API in a
dynamic and reproducible way.


```{r loadLibrary, eval=FALSE}
#install.packages("OECD")
library(OECD)
```

```{r}
dataset_list <- get_datasets()
search_dataset("pension", data = dataset_list)
```


## Probando datos del SERVEL 

El sitio del SERVEL tiene información histórica de las elecciones

lea: SERVEL/readme.md

```{r}
#getwd()
#SelWorkingDir <- choose.dir()
#SelWorkingDir
file1<-"SERVEL\\VW_VOTARON_2017_1V.csv"
VW_VOTARON_2017_1V<- read.csv(file=file1, header=TRUE, sep=";")
head(VW_VOTARON_2017_1V)
```

